### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part2.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Ontologies}
\label{sec:ontologies}
This section provides a brief introduction to the concept of ontologies.

\subsection{Definitions}
As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
set of representational primitives with which to model a domain of
knowledge or discourse{\textquotedblright}. These primitives are
typically classes, attributes and relations among classes or class
instances (and others, such as function terms, rules, restrictions,
axioms and events). Their definitions provide information about their
semantics (meaning) and constraints on their logically consistent
usage.

@@

\subsection{General Description}
An ontology is used to define the vocabulary that agents use to exchange
queries and assertions. If an agent commits to a common ontology, it
guarantees the consistency (but not completeness) of the queries and
assertions using the vocabulary defined in the ontology \cite{gruber93towards}.
An agent supporting the interface defined in an ontology is not
required to use the terms of the ontology as an internal encoding of
its knowledge. Nonetheless, the definitions and formal constraints of
the ontology do restrict what can be meaningfully stated in the
language defined.

@@

\section{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:mediationproblem}
Ontologies are employed as explicit descriptions of the information
source semantics. According to \cite{wache2001ontology_information_integration}, integrating these
semantics along multiple systems can be done using three main
approaches. (1) The single ontology approach, using a single ontology
as a global shared vocabulary. Unfortunately this can not provide
different views on a domain, and using it for multiple similar
catalogues is unfeasible due to the vast amount of information needed
to be captured, and the susceptibility of the conceptualisation to
change. (2) The multiple ontology approach has each information source
describing its own ontology. This solves problems related to the single
ontology approach, but introduces the problem of differences in
conceptualising the same concepts. Ontology mediation can be used to
map different parts of the ontologies to each other. (3) The hybrid
approach uses multiple ontologies, but uses a global ontology to make
all the ontologies comparable to each other. This is closely related to
ontology merging, where overlapping entities are merged, and the rest
is simply composed into one unified ontology, allowing for the
definition of the concepts of both ontologies. We suggest that the FAST
Gadget Ontology (as defined in D2.1) could act as a global ontology in
the context of FAST, although further research needs to be done in this
direction, since the Gadget Ontology has a slightly different
epistemological commitment.

@@

\subsection{Motivating Example}
\label{subsec:mediationproblem_motivatingexample}
This section presents a brief example, illustrating a problem that
presents the need for ontology mediation.

@@

\subsection{Problem Statement}
\label{subsec:mediationproblem_problemstatement}
Given two ontologies \texttt{o} and
\texttt{o{\textquotesingle}} that need to be mapped to
each other, according to \cite{shvaiko2005schema_based} we can define a
mapping element as a 5-tuple \texttt{{\textless}id, e,
e{\textquotesingle}, n, R{\textgreater}}, where

@@

\subsection{Mediation Examples}
\label{subsec:mediationproblem_examples}
We present two further examples to better illustrate the use of ontology
mediation.

@@

\section{Ontology Mismatches}
\label{sec:ontologymismatches}
Ontology mismatches are the main obstacles in the combined use of
independently developed ontologies. We will now give an overview of the
main type of ontology mismatches as presented in \cite{klein2001combiningOntologies}, which
is the main source of input to this state-of-the-art report. We will
classify ontology mismatches into given types and will find their
relationships.

@@

\subsection{Language Level Mismatches}
\label{subsec:ontologymismatches_languagelevel}
We define four types of mismatches, although they often coincide. These
occur when ontologies written in different languages are combined. We
note that these type of mismatches will be less relevant for FAST
because we assume that standards for ontology languages are now so well
established that in practice we will not very often find ontologies
written in different languages anymore, but we discuss them briefly to
give a better overview of the meditation problem.

\subsubsection{Syntax}
Different languages obviously use different syntax. Therefore the same
concept can (and will be) often modeled in a completely different way.
For example, the concept of chairs would be defined in RDF Schema as
\texttt{{\textless}rdfs:Class
ID={\textquotedblright}Chair{\textquotedblright}{\textgreater}}, while
in LOOM it would be \texttt{(defconcept Chair)}.

@@

\subsubsection{Logical representation}
A slightly more complicated mismatch is when a logical notion can not be
represented in a given language. For example, in certain languages it
is possible to assert that two classes are disjoint, say using
\texttt{(disjoint A B),} whereas in other languages it is
necessary to express this using simpler statements, such as the
negation of set-inclusion statements, like \texttt{A
subclass-of (NOT B), B subclass-of (NOT A).}

@@

\subsubsection{Semantics of primitives}
This difference is a more subtle one, and occurs at the meta-model of
the language. It sometimes occurs that the semantics of an identical
name differs from language to language. For example, the expression
\texttt{A equalsTo B} can have many different
interpretations.

@@

\subsubsection{Language expressiveness}
This mismatch, at the meta-model level of the languages, is the one
having the greatest impact on the attempt to combine and reconcile
ontologies.

@@

\subsection{Ontology Level Mismatches}
\label{subsec:ontologymismatches_ontologylevel}
Mismatches at the model level happen both in the case of having the two
ontologies written in the same language and syntax and in the case of
different languages. We can distinguish between several types of
ontology level mismatches.

A conceptualisation mismatch is a difference in how the domain is
modeled (conceptualised), therefore we will have different concepts and
different relations between them. Explication mismatches are
differences in the way the concepts are specified. These occur as
mismatches in definitions, or terms, or the combination of the two.
Terminological mismatches are naming differences or conflicts. Finally,
encoding mismatches are differences in the way values are specified,
for example currencies for prices.

@@

\subsubsection{Conceptualisation Mismatches}
These are differences in the conceptualisation of the domain, that is
not only the specification differs in the two ontologies, but the model
as well.

@@

\subsubsection{Explication Mismatches}
The first two types of explication mismatches (\textit{paradigm} and
\textit{concept description}) result from the \textbf{style of
modeling.} The next two types (\textit{synonym terms} and
\textit{homonym terms}) can be classified as \textbf{terminological
mismatches}. Finally, there is one trivial difference, cause by
\textit{encoding} mismatches.

@@

\section{Ontology Mediation Approaches}
\label{sec:mediationapproaches}
In this section we give an overview of some of the major approaches to
ontology mediation, following the structure of Section 2: Approaches in
Ontology Mediation of \cite{debruin2005wsml}. Firstly, in Section
\ref{subsec:mediationapproaches_ontologymapping} we present some representative approaches to the specification of
mappings. Then in Section \ref{subsec:mediationapproaches_ontologyalignment} we overview some of the main approaches
and algorithms to overcome mismatches and find mappings between
ontologies, that is we discuss the topic of ontology alignment. Finally
in Section \ref{subsec:mediationapproaches_ontologymerging} we discuss the process of merging several ontologies.
Further research is needed to decide on the type of ontology mediation
to be used in the context of FAST, the algorithms and approaches
applied, and whether to use alignment or merging.

\subsection{Ontology Mapping}
\label{subsec:mediationapproaches_ontologymapping}
An ontology mapping is a declarative specification of the semantic
overlap between two ontologies \cite{debruin2005wsml}. It is the
result of the ontology alignment process (see Figure 3.). This mapping
is usually expressed as a set of axioms in a certain mapping language.

@@

\begin{itemize}
\item \texttt{activity} , which is performed during a
certain time interval; 
\item \texttt{activity} \texttt{occurrence} ,
contains the snapshot of an activity at a moment in time; 
\item \texttt{timepoint} , marks the time interval of the
activity; 
\item \texttt{objects} , things that do not possess
temporal properties, but may participate in certain activities at given
timepoints. 
\end{itemize}
\subparagraph{RDFT.}
The basic RDFT class is \texttt{Bridge }that connects two
concepts. It describes common properties of bridges, which enables one
to specify correspondences between an entity and a set of entities.
allowing for one-to-many and many-to-one connections.

@@

\subsection{Ontology Alignment}
\label{subsec:mediationapproaches_ontologyalignment}
\subsubsection{The Match Operator}
Ontology alignment is the process of discovering similarities between
two source ontologies \cite{debruin2005wsml}. This process can be
described as the application of the so-called \textit{Match} operator
described in \cite{rahm01survey}. The result of the matching
process is a specification of the similarities between the two
ontologies. The input of the Match operator consists mainly of two
ontologies and some other inputs (initial mapping that will be
extended, mapping parameters that configure the process and external
resources used by the matching process; see Section 3.2).

@@

\subsubsection{The Alignment Process}
The alignment process proposed in \cite{debruin2005wsml} relieves the user of
some of the burdens in creating the mappings. The input of the process
consists of two ontologies which are to be aligned. The output is a set
of mappings.

@@

\subsubsection{Classification of Approaches}
An implementation of Match may use multiple matching algorithms or
\textit{matchers. }This allows us to select matchers according to the
application domain or schema types.

@@

\subsubsection{Overview of Approaches}
In this section we give a brief overview of some of the main approaches
to ontology alignment.

@@

The output of one iteration of QOM can be used as part of the input of a
subsequent iteration in order to refine the result. After a number of
iterations the actual mapping between the ontologies is obtained. 
\subsubsection{Alignment Examples}
This section provides a few examples of the matching
process{\textquotesingle} results.
\begin{center}
\tablecaption{Constraint-based matching example. {\newline}Taken from \cite{rahm01survey}.}
\label{tab:ConstraintMatching}
\tablehead{}
%begin{supertabular}{|r@{\hspace{6.5mm}}|r@{\hspace{5.5mm}}|r|r|}
\begin{supertabular}{|m{0.4\textwidth}|m{0.4\textwidth}|}
\hline
S1 elements &
S2 elements\\\hline
Employee

@@

\subsection{Ontology Merging}
\label{subsec:mediationapproaches_ontologymerging}
Ontology merging is the creation of a new ontology from one or more
source ontologies. The new ontology will unify and in general replace
the original source ontologies.

@@

\subsubsection{PROMPT}
\label{subsubsec:mediationapproaches_ontologymerging_PROMPT}
PROMPT \cite{noy2000prompt} is an algorithm and interactive tool for
ontology merging and alignment. It was one of the first ontology
merging tools, comparing every pair in the two ontologies and looking
for syntactic similarity between labels. It has been implemented as a
Proteg\'e-2000 extension.

@@

\subsubsection{OntoMerge}
OntoMerge \cite{dou2002ontology_translation} is an online ontology translation system,
which translates a dataset to a new dataset which captures the same
information in a different ontology.

>>>>

\section{Ontology Mediation Background}
\label{sec:background}

\subsection{Ontologies}
\label{sec:ontologies}
This section provides a brief introduction to the concept of ontologies.

\subsubsection{Definitions}
As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
set of representational primitives with which to model a domain of
knowledge or discourse{\textquotedblright}. These primitives are
typically classes, attributes and relations among classes or class
instances (and others, such as function terms, rules, restrictions,
axioms and events). Their definitions provide information about their
semantics (meaning) and constraints on their logically consistent
usage.

@@

\subsubsection{General Description}
An ontology is used to define the vocabulary that agents use to exchange
queries and assertions. If an agent commits to a common ontology, it
guarantees the consistency (but not completeness) of the queries and
assertions using the vocabulary defined in the ontology \cite{gruber93towards}.
An agent supporting the interface defined in an ontology is not
required to use the terms of the ontology as an internal encoding of
its knowledge. Nonetheless, the definitions and formal constraints of
the ontology do restrict what can be meaningfully stated in the
language defined.

@@

\subsection{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:mediationproblem}
Ontologies are employed as explicit descriptions of the information
source semantics. According to \cite{wache2001ontology_information_integration}, integrating these
semantics along multiple systems can be done using three main
approaches. (1) The single ontology approach, using a single ontology
as a global shared vocabulary. Unfortunately this can not provide
different views on a domain, and using it for multiple similar
catalogues is unfeasible due to the vast amount of information needed
to be captured, and the susceptibility of the conceptualisation to
change. (2) The multiple ontology approach has each information source
describing its own ontology. This solves problems related to the single
ontology approach, but introduces the problem of differences in
conceptualising the same concepts. Ontology mediation can be used to
map different parts of the ontologies to each other. (3) The hybrid
approach uses multiple ontologies, but uses a global ontology to make
all the ontologies comparable to each other. This is closely related to
ontology merging, where overlapping entities are merged, and the rest
is simply composed into one unified ontology, allowing for the
definition of the concepts of both ontologies. We suggest that the FAST
Gadget Ontology (as defined in D2.1) could act as a global ontology in
the context of FAST, although further research needs to be done in this
direction, since the Gadget Ontology has a slightly different
epistemological commitment.

@@

\subsubsection{Motivating Example}
\label{subsec:mediationproblem_motivatingexample}
This section presents a brief example, illustrating a problem that
presents the need for ontology mediation.

@@

\subsubsection{Problem Statement}
\label{subsec:mediationproblem_problemstatement}
Given two ontologies \texttt{o} and
\texttt{o{\textquotesingle}} that need to be mapped to
each other, according to \cite{shvaiko2005schema_based} we can define a
mapping element as a 5-tuple \texttt{{\textless}id, e,
e{\textquotesingle}, n, R{\textgreater}}, where

@@

\subsubsection{Mediation Examples}
\label{subsec:mediationproblem_examples}
We present two further examples to better illustrate the use of ontology
mediation.

@@

\subsection{Ontology Mismatches}
\label{sec:ontologymismatches}
Ontology mismatches are the main obstacles in the combined use of
independently developed ontologies. We will now give an overview of the
main type of ontology mismatches as presented in \cite{klein2001combiningOntologies}, which
is the main source of input to this state-of-the-art report. We will
classify ontology mismatches into given types and will find their
relationships.

@@

\subsubsection{Language Level Mismatches}
\label{subsec:ontologymismatches_languagelevel}
We define four types of mismatches, although they often coincide. These
occur when ontologies written in different languages are combined. We
note that these type of mismatches will be less relevant for FAST
because we assume that standards for ontology languages are now so well
established that in practice we will not very often find ontologies
written in different languages anymore, but we discuss them briefly to
give a better overview of the meditation problem.

\paragraph{Syntax}
Different languages obviously use different syntax. Therefore the same
concept can (and will be) often modeled in a completely different way.
For example, the concept of chairs would be defined in RDF Schema as
\texttt{{\textless}rdfs:Class
ID={\textquotedblright}Chair{\textquotedblright}{\textgreater}}, while
in LOOM it would be \texttt{(defconcept Chair)}.

@@

\paragraph{Logical representation}
A slightly more complicated mismatch is when a logical notion can not be
represented in a given language. For example, in certain languages it
is possible to assert that two classes are disjoint, say using
\texttt{(disjoint A B),} whereas in other languages it is
necessary to express this using simpler statements, such as the
negation of set-inclusion statements, like \texttt{A
subclass-of (NOT B), B subclass-of (NOT A).}

@@

\paragraph{Semantics of primitives}
This difference is a more subtle one, and occurs at the meta-model of
the language. It sometimes occurs that the semantics of an identical
name differs from language to language. For example, the expression
\texttt{A equalsTo B} can have many different
interpretations.

@@

\paragraph{Language expressiveness}
This mismatch, at the meta-model level of the languages, is the one
having the greatest impact on the attempt to combine and reconcile
ontologies.

@@

\subsubsection{Ontology Level Mismatches}
\label{subsec:ontologymismatches_ontologylevel}
Mismatches at the model level happen both in the case of having the two
ontologies written in the same language and syntax and in the case of
different languages. We can distinguish between several types of
ontology level mismatches.

A conceptualisation mismatch is a difference in how the domain is
modelled (conceptualised), therefore we will have different concepts and
different relations between them. Explication mismatches are
differences in the way the concepts are specified. These occur as
mismatches in definitions, or terms, or the combination of the two.
Terminological mismatches are naming differences or conflicts. Finally,
encoding mismatches are differences in the way values are specified,
for example currencies for prices.

@@

\paragraph{Conceptualisation Mismatches}
These are differences in the conceptualisation of the domain, that is
not only the specification differs in the two ontologies, but the model
as well.

@@

\paragraph{Explication Mismatches}
The first two types of explication mismatches (\textit{paradigm} and
\textit{concept description}) result from the \textbf{style of
modelling.} The next two types (\textit{synonym terms} and
\textit{homonym terms}) can be classified as \textbf{terminological
mismatches}. Finally, there is one trivial difference, cause by
\textit{encoding} mismatches.

@@

\subsection{Ontology Mediation Approaches}
\label{sec:mediationapproaches}
In this section we give an overview of some of the major approaches to
ontology mediation, following the structure of Section 2: Approaches in
Ontology Mediation of \cite{debruin2005wsml}. Firstly, in Section
\ref{subsec:mediationapproaches_ontologymapping} we present some representative approaches to the specification of
mappings. Then in Section \ref{subsec:mediationapproaches_ontologyalignment} we overview some of the main approaches
and algorithms to overcome mismatches and find mappings between
ontologies, that is we discuss the topic of ontology alignment. Finally
in Section \ref{subsec:mediationapproaches_ontologymerging} we discuss the process of merging several ontologies.
Further research is needed to decide on the type of ontology mediation
to be used in the context of FAST, the algorithms and approaches
applied, and whether to use alignment or merging.

\subsubsection{Ontology Mapping}
\label{subsec:mediationapproaches_ontologymapping}
An ontology mapping is a declarative specification of the semantic
overlap between two ontologies \cite{debruin2005wsml}. It is the
result of the ontology alignment process (see Figure 3.). This mapping
is usually expressed as a set of axioms in a certain mapping language.

@@

\begin{itemize}
\item \texttt{activity} , which is performed during a
certain time interval; 
\item \texttt{activity} \texttt{occurrence} ,
contains the snapshot of an activity at a moment in time; 
\item \texttt{timepoint} , marks the time interval of the
activity; 
\item \texttt{objects} , things that do not possess
temporal properties, but may participate in certain activities at given
time points. 
\end{itemize}
\subparagraph{RDFT.}
The basic RDFT class is \texttt{Bridge }that connects two
concepts. It describes common properties of bridges, which enables one
to specify correspondences between an entity and a set of entities.
allowing for one-to-many and many-to-one connections.

@@

\subsubsection{Ontology Alignment}
\label{subsec:mediationapproaches_ontologyalignment}
\subsubsection{The Match Operator}
Ontology alignment is the process of discovering similarities between
two source ontologies \cite{debruin2005wsml}. This process can be
described as the application of the so-called \textit{Match} operator
described in \cite{rahm01survey}. The result of the matching
process is a specification of the similarities between the two
ontologies. The input of the Match operator consists mainly of two
ontologies and some other inputs (initial mapping that will be
extended, mapping parameters that configure the process and external
resources used by the matching process; see Section 3.2).

@@

\paragraph{The Alignment Process}
The alignment process proposed in \cite{debruin2005wsml} relieves the user of
some of the burdens in creating the mappings. The input of the process
consists of two ontologies which are to be aligned. The output is a set
of mappings.

@@

\paragraph{Classification of Approaches}
An implementation of Match may use multiple matching algorithms or
\textit{matchers. }This allows us to select matchers according to the
application domain or schema types.

@@

\paragraph{Overview of Approaches}
In this section we give a brief overview of some of the main approaches
to ontology alignment.

@@

The output of one iteration of QOM can be used as part of the input of a
subsequent iteration in order to refine the result. After a number of
iterations the actual mapping between the ontologies is obtained. 
\paragraph{Alignment Examples}
This section provides a few examples of the matching
process{\textquotesingle} results.
\begin{center}
\tablecaption{Constraint-based matching example. {\newline}Taken from \cite{rahm01survey}.}
\label{tab:ConstraintMatching}
\tablehead{}
%begin{supertabular}{|r@{\hspace{6.5mm}}|r@{\hspace{5.5mm}}|r|r|}
\begin{supertabular}{|m{0.4\textwidth}|m{0.4\textwidth}|}
\hline
S1 elements &
S2 elements\\\hline
Employee

@@

\subsubsection{Ontology Merging}
\label{subsec:mediationapproaches_ontologymerging}
Ontology merging is the creation of a new ontology from one or more
source ontologies. The new ontology will unify and in general replace
the original source ontologies.

@@

\paragraph{PROMPT}
\label{subsubsec:mediationapproaches_ontologymerging_PROMPT}
PROMPT \cite{noy2000prompt} is an algorithm and interactive tool for
ontology merging and alignment. It was one of the first ontology
merging tools, comparing every pair in the two ontologies and looking
for syntactic similarity between labels. It has been implemented as a
Proteg\'e-2000 extension.

@@

\paragraph{OntoMerge}
OntoMerge \cite{dou2002ontology_translation} is an online ontology translation system,
which translates a dataset to a new dataset which captures the same
information in a different ontology.
